### 2021-11-07 publication/derivations.tex
neaten up appendix
Add impl details

\section{Appendix: Derivations}

@@

\subsection{Derivation of {\boldmath $\;U(\theta)$}}
\label{appdx:form-U}

Recall from~(\ref{eq:U}) in Section~\ref{s:sfb} that,
$$	\pi_\theta(\theta) \propto p(\theta | X, b) \propto p(b | X, \theta) p(\theta) \propto  \exp \left( - U(\theta) \right),
$$ 
so that $U$ can be expressed as,
$$
U(\theta) 
= - \left( \log p(b | X, \theta) + \log p(\theta) \right) + \textrm{const}.
$$
Writing,
$y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and 
$a_{ij} \coloneqq \phi_j(x_i; \theta)$, we have that,
%
\begin{equation}
	\log p(b | X, \theta) = \sum_{i \in [N]} \sum_{j \in [B]} y_{ij} \log a_{ij}  \quad \textrm{and} \quad
	\log p(\theta) = -\frac{D B}{2} \log 2\pi - \frac{1}{2 \sigma_\theta^2} 
	\|\theta \|^2,
	\label{eqn:U-constituent-terms}
\end{equation}
%
where
$\|\theta\|^2 = \sum_{i} \theta_{i}^2 = \sum_{j \in [B]} \|w_j\|^2$ 
is the Euclidean norm of the vector of parameters $\theta$.
Therefore, discarding constant terms, we 
obtain,
%
\begin{equation}
	U(\theta) = \left( \sum_{i \in [N]} \sum_{j \in [B]} y_{ij} \log \frac{1}{a_{ij}} \right)
	+ \frac{1}{2\sigma_\theta^2} \|\theta\|^2.
\end{equation}

\subsection{Derivation of {\boldmath $\;\nabla U(\theta)$}}
\label{appdx:gradu}
Here we show how the gradient 
$\nabla U(\theta)$ can be computed explicitly.
As
$\theta = \{ w_k \}_{k=1}^{B}$,
to compute the gradient
$\nabla U(\theta)$ we need to compute
each of its components,
$\nicefrac{\partial U}{\partial w_k}$,
$1\leq k\leq B$.
To that end, we first compute,
%
\begin{align}
	\frac{\partial a_{ij}}{\partial w_k} &= \frac
	{x_i \exp(w_j^T x_i) \delta_{jk} \cdot \sum_{r \in [B]} \exp(w_r^T x_i) 
		- 
		\exp(w_j^T x_i) \cdot x_i \exp(w_k^T x_i)}
	{\left( \sum_{r \in [B]} \exp(w_r^T x_i) \right)^2} \nonumber \\
	&= x_i \left( a_{ij} \delta_{jk} - a_{ij}a_{ik} \right), 
	\label{eq:dadw}
\end{align}
%
where $\delta_{jk} \coloneqq \one \left\{ j = k \right\}$,
and we also easily find,
%
\begin{equation}
	\frac{ \partial}{\partial w_k} \|\theta\|^2 = \frac{\partial}{\partial w_k} \left( \sum_{r \in [B]} \|w_r\|^2 \right) = 2w_k.
	\label{eq:dtsdw}
\end{equation}
%
Recall the expression for $U(\theta)$ in~(\ref{eqn:U-form}). Using~(\ref{eq:dadw}) and~(\ref{eq:dtsdw}), we obtain,
\begin{align}
	\frac{\partial U}{\partial w_k} &= 
	\sum_{i \in [N]} \sum_{j \in [B]} y_{ij} 
	\left( -\frac{x_i}{a_{ij}} \left( a_{ij} \delta_{jk} - a_{ij} a_{ik} \right) \right)
	+ \frac{w_k}{\sigma_\theta^2} \nonumber \\
	&= - \left( \sum_{i \in [N]} \Big\{ x_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right).
\end{align}
%
This can be computed 
efficiently through matrix operations. The only property of $y_{ij}$ 
we have used in the derivation is the constraint $\sum_{j \in [B]} y_{ij} = 1$,
for all $i$.

>>>>

\section{Appendix}

@@

\subsection{Derivation of {\boldmath $U(\theta)$} and {\boldmath $\nabla U(\theta)$}}
\label{appdx:form-U}

Recall from~(\ref{eq:U}) in Section~\ref{s:sfb} that,
$$	\pi_\theta(\theta) \propto p(\theta | X, b) \propto p(b | X, \theta) p(\theta) \propto  \exp \left( - U(\theta) \right),
$$ 
so that $U$ can be expressed as,
$$
U(\theta) 
= - \left( \log p(b | X, \theta) + \log p(\theta) \right) + \textrm{const}.
$$
Writing,
$y_{ij} \coloneqq \one \left\{ b_i = j \right\}$ and 
$a_{ij} \coloneqq \phi_j(x_i; \theta)$, we have that,
%
\begin{equation*}
	\log p(b | X, \theta) = \sum_{i \in [N]} \sum_{j \in [B]} y_{ij} \log a_{ij}  \quad \textrm{and} \quad
	\log p(\theta) = -\frac{D B}{2} \log 2\pi - \frac{1}{2 \sigma_\theta^2} 
	\|\theta \|^2,
	\label{eqn:U-constituent-terms}
\end{equation*}
%
where
$\|\theta\|^2 = \sum_{i} \theta_{i}^2 = \sum_{j \in [B]} \|w_j\|^2$ 
is the Euclidean norm of the vector of parameters $\theta$.
Therefore, discarding constant terms, we 
obtain,
%
\begin{equation}
	U(\theta) = \left( \sum_{i \in [N]} \sum_{j \in [B]} y_{ij} \log \frac{1}{a_{ij}} \right)
	+ \frac{1}{2\sigma_\theta^2} \|\theta\|^2.
\end{equation}

As
$\theta = \{ w_k \}_{k=1}^{B}$,
to compute the gradient
$\nabla U(\theta)$ we need to compute
each of its components,
$\nicefrac{\partial U}{\partial w_k}$,
$1\leq k\leq B$.
To that end, we first compute,
%
\begin{align}
	\frac{\partial a_{ij}}{\partial w_k} &= \frac
	{x_i \exp(w_j^T x_i) \delta_{jk} \cdot \sum_{r \in [B]} \exp(w_r^T x_i) 
		- 
		\exp(w_j^T x_i) \cdot x_i \exp(w_k^T x_i)}
	{\left( \sum_{r \in [B]} \exp(w_r^T x_i) \right)^2} \nonumber \\
	&= x_i \left( a_{ij} \delta_{jk} - a_{ij}a_{ik} \right), 
	\label{eq:dadw}
\end{align}
%
where $\delta_{jk} \coloneqq \one \left\{ j = k \right\}$,
and we also easily find,
%
\begin{equation}
	\frac{ \partial}{\partial w_k} \|\theta\|^2 = \frac{\partial}{\partial w_k} \left( \sum_{r \in [B]} \|w_r\|^2 \right) = 2w_k.
	\label{eq:dtsdw}
\end{equation}
%
Recall the expression for $U(\theta)$ in~(\ref{eqn:U-form}). Using~(\ref{eq:dadw}) and~(\ref{eq:dtsdw}), we obtain,
\begin{align}
	\frac{\partial U}{\partial w_k} &= 
	\sum_{i \in [N]} \sum_{j \in [B]} y_{ij} 
	\left( -\frac{x_i}{a_{ij}} \left( a_{ij} \delta_{jk} - a_{ij} a_{ik} \right) \right)
	+ \frac{w_k}{\sigma_\theta^2} \nonumber \\
	&= - \left( \sum_{i \in [N]} \Big\{ x_i (y_{ik} - a_{ik}) \Big\} - \frac{w_k}{\sigma_\theta^2} \right).
\end{align}
%
This can be computed 
efficiently through matrix operations. The only property of $y_{ij}$ 
we have used in the derivation is the constraint $\sum_{j \in [B]} y_{ij} = 1$,
for all $i$.

\subsection{Implementation details}
\label{appdx:hyperparams}

All data analysis and visualisation was implemented in Python. Full source code is available at:
\url{https://github.com/LozzaTray/Jormungandr-code}. Table~\ref{tab:hyperparams} contains the hyper-parameters for each experiment run. The thinned samples are generated as,
\begin{equation}
	\Tcal_\star = \{T_\star \kappa_\star + i \lambda_\star :  
	0 \leq i \leq \lfloor T_\star(1 - \kappa_\star) / \lambda_\star \rfloor \}.
\end{equation}

\begin{table}[!h]
	\centering
	\caption{Hyper-parameter values for each experiment}
	\label{tab:hyperparams}
	\resizebox{\textwidth}{!}{%
		\begin{tabular}{c|ccc|ccc|cccc|cc|cccc}
			Dataset & 
			$B$ & $f$ & $\sigma_\theta$ & 
			$T_b$ & $\kappa_b$ & $\lambda_b$ & 
			$T_\theta$ & $\kappa_\theta$ & $\lambda_\theta$ & $s$ &
			$k$ & $D'$ &
			$T_\theta'$ & $\kappa_\theta'$ & $\lambda_\theta'$ & $s'$
			\\ \hline
			Polbooks &
			3 & 0.7 & 1 &
			1,000 & 0.2 & 5 &
			10,000 & 0.4 & 10 & 0.05 &
			-- & -- & 
			-- & -- & -- & -- \\
			School &
			10 & 0.7 & 1 &
			1,000 & 0.2 & 5 &
			10,000 & 0.4 & 10 & 0.2 &
			1 & 10 & 
			10,000 & 0.4 & 10 & 0.2 \\
			FB Egonet &
			10 & 0.7 & 1 &
			1,000 & 0.2 & 5 &
			10,000 & 0.4 & 10 & 0.017 &
			1 & 10 & 
			10,000 & 0.4 & 10 & 0.5 \\
		\end{tabular}
	}
\end{table}

### 2021-11-07 publication/experiments.tex
Add impl details

Next, to assess the performance of the feature-to-block predictor, 
we randomly partition the vertex set $[N]$ into training and test sets: $\Gcal_0$ and $\Gcal_1$.
\footnote{We choose $f=0.7$ as the fraction of vertices that go to form our test set. Please refer to \url{https://github.com/LozzaTray/Jormungandr-code} for full source code.}
The $b$-chain is run using the whole network but we only use vertices $v \in \Gcal_0$ to train the $\theta$-chain. As $|\Gcal_0| \neq |\Gcal_1|$ in general, we use the average cross-entropy loss 
over each set to gauge fit,
%
\begin{equation}
	\bar{\Lcal}_\star \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \Lcal_\star^{(t)},
	\quad \textrm{where} \quad
	\Lcal_\star^{(t)} \coloneqq \frac{1}{|\Gcal_\star|} \sum_{i \in \Gcal_\star}\sum_{j \in [B]} \hat{y}_{ij} \log \frac{1}{\phi_j \left(x_i; \theta^{(t)} \right)},
	\label{eqn:cross-entropy-loss}
\end{equation}
%
where $\star \in \{0, 1\}$ toggles between the training and test sets.
%
Nevertheless, the cross-entropy loss is a coarse measure of fit. We wish to define a new measure of fit specific to each detected block. Let us define
$
	\Bcal_\star(j) \coloneqq \{i \in \Gcal_\star : \hat{b}_i = j\}
$
where
$ 
	\hat{b}_i \coloneqq \argmax_j \hat{y}_{ij}.
$
Now $\Bcal_\star(j)$ is the set of vertices that have maximum a posteriori probability of belonging to block $j$. We now define the {\em block-accuracy} for block $j$ as in \ref{eqn:accuracy}. This effectively tests whether the feature-to-block and the graph-to-block predictions agree in their largest component.
%
\begin{equation}
	\eta_\star(j) \coloneqq \frac{1}{|\Bcal_\star (j)| \cdot 
	|\Tcal_\theta| } 
	\sum_{i \in \Bcal_\star (j)}  \sum_{t \in \Tcal_\theta}
	\one \left\{\hat{b}_i = \argmax_j \phi_j \left( x_i; \theta^{(t)} \right) \right\}.
	\label{eqn:accuracy}
\end{equation}

>>>>

Next, to assess the performance of the feature-to-block predictor, 
we randomly partition the vertex set $[N]$ such that a constant fraction $f$ forms our training set $\Gcal_0$ and the remainder form our test set $\Gcal_1$.
The $b$-chain is run using the whole network but we only use vertices $v \in \Gcal_0$ to train the $\theta$-chain. As $|\Gcal_0| \neq |\Gcal_1|$ in general, we use the average cross-entropy loss 
over each set to gauge fit,
%
\begin{equation}
	\bar{\Lcal}_\star \coloneqq \frac{1}{|\Tcal_\theta|} \sum_{t \in \Tcal_\theta} \Lcal_\star^{(t)},
	\quad \textrm{where} \quad
	\Lcal_\star^{(t)} \coloneqq \frac{1}{|\Gcal_\star|} \sum_{i \in \Gcal_\star}\sum_{j \in [B]} \hat{y}_{ij} \log \frac{1}{\phi_j \left(x_i; \theta^{(t)} \right)},
	\label{eqn:cross-entropy-loss}
\end{equation}
%
where $\star \in \{0, 1\}$ toggles between the training and test sets.
%
Nevertheless, the cross-entropy loss is a coarse measure of fit. We wish to define a new measure of fit specific to each detected block. Let us define
$
	\Bcal_\star(j) \coloneqq \{i \in \Gcal_\star : \hat{b}_i = j\}
$
where
$ 
	\hat{b}_i \coloneqq \argmax_j \hat{y}_{ij}.
$
Now $\Bcal_\star(j)$ is the set of vertices that have maximum a posteriori probability of belonging to block $j$. We now define the {\em block-accuracy} for block $j$ as in \ref{eqn:accuracy}. This effectively tests whether the feature-to-block and the graph-to-block predictions agree in their largest component.
%
\begin{equation}
	\eta_\star(j) \coloneqq \frac{1}{|\Bcal_\star (j)| \cdot 
	|\Tcal_\theta| } 
	\sum_{i \in \Bcal_\star (j)}  \sum_{t \in \Tcal_\theta}
	\one \left\{\hat{b}_i = \argmax_j \phi_j \left( x_i; \theta^{(t)} \right) \right\}.
	\label{eqn:accuracy}
\end{equation}
